Ethereum: A Secure Decentralised Generalised Transaction Ledger \\ \smaller \textbf{{Istanbul version \YellowPaperVersionNumber}}

### 2021-06-28 Paper.tex
3607: update to use empty code hash instead of codesize

Sometimes, a path follows a new protocol from a particular height (block number).
This document describes one version of the protocol, namely the \textit{Istanbul} version defined in EIP-1679 by \cite{EIP-1679}.
In order to follow back the history of a path, one must reference multiple versions of this document.
Here are the block numbers of protocol updates on the Ethereum main network:
\par
\begin{center}
\begin{tabular}{lr}
\toprule
Name & First Block Number \\
\midrule
$F_{\mathrm{Homestead}}$         &  1150000 \\
$F_{\mathrm{Tangerine Whistle}}$ &  2463000 \\
$F_{\mathrm{Spurious Dragon}}$   &  2675000 \\
$F_{\mathrm{Byzantium}}$         &  4370000 \\
$F_{\mathrm{Constantinople}}$    &  7280000 \\
$F_{\mathrm{Petersburg}}$        &  7280000 \\
$F_{\mathrm{Istanbul}}$          &  9069000 \\
$F_{\mathrm{Muir Glacier}}$      &  9200000 \\
$F_{\mathrm{Berlin}}$            & 12244000 \\
\bottomrule
\end{tabular}
\end{center}
\par

Occasionally actors do not agree on a protocol change, and a permanent fork occurs.
In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
For the Ethereum main network
\begin{equation}
  \linkdest{chain_id}
  \beta = 1
\end{equation}

@@

Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
\begin{equation}
\texttt{TRIE}\big(L_{\mathrm{I}}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
\end{equation}

The collapse function for the set of key/value pairs in the trie, $L_{\mathrm{I}}^*$, is defined as the element-wise transformation of the base function $L_{\mathrm{I}}$, given as:
\begin{equation}
L_{\mathrm{I}}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
\end{equation}

@@

Thus we may define a world-state collapse function $L_{\mathrm{S}}$:
\begin{equation}
L_{\mathrm{S}}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
\end{equation}
where
\begin{equation}
p(a) \equiv  \big(\texttt{KEC}(a), \texttt{RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
\end{equation}

This function, $L_{\mathrm{S}}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
\begin{equation}
\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
\end{equation}
\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
\begin{equation}
\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{N}_{256} \wedge x_{\mathrm{b}} \in \mathbb{N}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
\end{equation}

@@

A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum. While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents. \eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}. There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation'). Both types specify a number of common fields:

@@

\begin{equation}
L_{\mathrm{T}}(T) \equiv \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
\end{cases}
\end{equation}

@@

\begin{description}
\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
\end{description}
\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_{\mathrm{H}}, B_{\mathbf{T}}, B_{\mathbf{U}})
\end{equation}

@@

%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{\mathrm{B}}$; it is merely a convenience equivalence.

@@

\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
\begin{equation}
\begin{array}[t]{lclc}
\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{\mathrm{T}}(B_{\mathbf{T}}[i]))\}) & \wedge \\
\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, B_{\mathbf{R}}[i])\}) & \wedge \\
\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
\end{array}
\end{equation}
where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
\begin{equation}
p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
\end{equation}

Furthermore:
\begin{equation}
\mathtt{TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
\end{equation}

Thus $\texttt{TRIE}(L_{\mathrm{S}}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{\mathrm{H}})$ is the parent block of $B$, defined directly.

@@

\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{\mathrm{B}}$ and $L_{\mathrm{H}}$ are the preparation functions for a block and block header respectively.
We assert the types and order of the structure for when the RLP transformation is required:
\begin{eqnarray}
\quad L_{\mathrm{H}}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
\quad L_{\mathrm{B}}(B) & \equiv & \big( L_{\mathrm{H}}(B_{\mathrm{H}}), L_{\mathrm{T}}^*(B_{\mathbf{T}}), L_{\mathrm{H}}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
\end{eqnarray}

@@

We define $P(B_{\mathrm{H}})$ to be the parent block of $B$, formally:
\begin{equation}
P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_{\mathrm{H}})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
\end{equation}

\hypertarget{block_number_H__i}{}The block number is the parent's block number incremented by one:
\begin{equation}
H_{\mathrm{i}} \equiv {{P(H)_{\mathrm{H}}}_{\mathrm{i}}} + 1
\end{equation}

@@

\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
\begin{equation}
D(H) \equiv \begin{dcases}
\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
\text{max}\!\left(\mindifficulty, {P(H)_{\mathrm{H}}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
\end{dcases}
\end{equation}
where:
\begin{equation}
\mindifficulty \equiv 131072
\end{equation}
\begin{equation}
\diffadjustment \equiv \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{d}}}{2048}\right\rfloor
\end{equation}
\begin{equation}
\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{\mathrm{H}}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
\end{equation}
\begin{equation*}
y \equiv \begin{cases}
1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
2 & \text{otherwise}
\end{cases}
\end{equation*}
\begin{align}
\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - \kappa, 0) \\
\kappa &\equiv \begin{cases} 
  3000000  & \text{if} \quad F_{\mathrm{Byzantium}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Constantinople}} \\
  5000000  & \text{if} \quad F_{\mathrm{Constantinople}} \leqslant H_{\mathrm{i}} < F_{\mathrm{Muir Glacier}} \\
  9000000  & \text{if} \quad H_{\mathrm{i}} \geqslant F_{\mathrm{Muir Glacier}} \\
\end{cases}
\end{align}

@@

\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
\begin{eqnarray}
& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
\nonumber& & H_{\mathrm{l}} \geqslant 5000
\end{eqnarray}

\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
\begin{equation}
H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}}
\end{equation}

@@

\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
\begin{eqnarray}
V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} < {P(H)_{\mathrm{H}}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} > {P(H)_{\mathrm{H}}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{\mathrm{H}}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
\nonumber& & H_{\mathrm{s}} > {P(H)_{\mathrm{H}}}_{\mathrm{s}} \quad \wedge \\
\nonumber& & H_{\mathrm{i}} = {P(H)_{\mathrm{H}}}_{\mathrm{i}} +1 \quad \wedge \\
\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
\end{eqnarray}
where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$

@@

\begin{enumerate}
\item The transaction is well-formed RLP, with no additional trailing bytes;
\item the transaction signature is valid;
\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
\end{enumerate}

@@

After the message call or contract creation is processed, the refund counter has to be incremented for the accounts that were self-destructed throughout its invocation.
\begin{equation}
	A'_{\mathrm{r}} \equiv A_{\mathrm{r}} + \sum_{i \in A_{\mathbf{s}}} R_{\mathrm{selfdestruct}}
\end{equation}

@@

These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{sigma_n}{later} for state and nonce validation.

@@

$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{\mathrm{H}}$ has no special treatment and is determined from the blockchain.

>>>>

Sometimes, a path follows a new protocol from a particular height (block number).
This document describes one version of the protocol, namely the \textit{Istanbul} version defined in EIP-1679 by \cite{EIP-1679}.
In order to follow back the history of a path, one must reference multiple versions of this document.

Occasionally actors cannot agree on a protocol change, and a permanent fork occurs.
In order to distinguish between diverged blockchains, EIP-155 by \cite{EIP-155} introduced the concept of chain ID, which we denote by $\beta$.
For the Ethereum main network
\begin{equation}
  \linkdest{chain_id}
  \beta = 1
\end{equation}

@@

Since we typically wish to refer not to the trie's root hash but to the underlying set of key/value pairs stored within, we define a convenient equivalence:
\begin{equation}
\texttt{TRIE}\big(L_{I}^*(\boldsymbol{\sigma}[a]_{\mathbf{s}})\big) \equiv \boldsymbol{\sigma}[a]_{\mathrm{s}}
\end{equation}

The collapse function for the set of key/value pairs in the trie, $L_{I}^*$, is defined as the element-wise transformation of the base function $L_{I}$, given as:
\begin{equation}
L_{I}\big( (k, v) \big) \equiv \big(\texttt{KEC}(k), \texttt{RLP}(v)\big)
\end{equation}

@@

Thus we may define a world-state collapse function $L_{S}$:
\begin{equation}
L_{S}(\boldsymbol{\sigma}) \equiv \{ p(a): \boldsymbol{\sigma}[a] \neq \varnothing \}
\end{equation}
where
\begin{equation}
p(a) \equiv  \big(\texttt{KEC}(a), \texttt{RLP}\big( (\boldsymbol{\sigma}[a]_{\mathrm{n}}, \boldsymbol{\sigma}[a]_{\mathrm{b}}, \boldsymbol{\sigma}[a]_{\mathrm{s}}, \boldsymbol{\sigma}[a]_{\mathrm{c}}) \big) \big)
\end{equation}

This function, $L_{S}$, is used alongside the trie function to provide a short identity (hash) of the world state. We assume:
\begin{equation}
\forall a: \boldsymbol{\sigma}[a] = \varnothing \; \vee \; (a \in \mathbb{B}_{20} \; \wedge \; v(\boldsymbol{\sigma}[a]))
\end{equation}
\linkdest{account_validity_function_v__x}{}where $v$ is the account validity function:
\begin{equation}
\quad v(x) \equiv x_{\mathrm{n}} \in \mathbb{N}_{256} \wedge x_{\mathrm{b}} \in \mathbb{N}_{256} \wedge x_{\mathrm{s}} \in \mathbb{B}_{32} \wedge x_{\mathrm{c}} \in \mathbb{B}_{32}
\end{equation}

@@

A transaction (formally, $T$) is a single cryptographically-signed instruction constructed by an actor externally to the scope of Ethereum. The sender of a transaction can not be a contract. While it is assumed that the ultimate external actor will be human in nature, software tools will be used in its construction and dissemination\footnote{Notably, such `tools' could ultimately become so causally removed from their human-based initiation---or humans may become so causally-neutral---that there could be a point at which they rightly be considered autonomous agents. \eg contracts may offer bounties to humans for being sent transactions to initiate their execution.}. There are two types of transactions: those which result in message calls and those which result in the creation of new accounts with associated code (known informally as `contract creation'). Both types specify a number of common fields:

@@

\begin{equation}
L_{T}(T) \equiv \begin{cases}
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{i}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{if} \; T_{\mathrm{t}} = \varnothing\\
(T_{\mathrm{n}}, T_{\mathrm{p}}, T_{\mathrm{g}}, T_{\mathrm{t}}, T_{\mathrm{v}}, T_{\mathbf{d}}, T_{\mathrm{w}}, T_{\mathrm{r}}, T_{\mathrm{s}}) & \text{otherwise}
\end{cases}
\end{equation}

@@

\begin{description}
\item[parentHash]\linkdest{parent_Hash_H__p_def_words}{} The Keccak 256-bit hash of the parent block's header, in its entirety; formally $H_{\mathrm{p}}$.
\item[ommersHash] The Keccak 256-bit hash of the ommers list portion of this block; formally $H_{\mathrm{o}}$.
\item[beneficiary]\linkdest{beneficiary_H__c}{}\linkdest{H__c} The 160-bit address to which all fees collected from the successful mining of this block be transferred; formally $H_{\mathrm{c}}$.
\item[stateRoot] The Keccak 256-bit hash of the root node of the state trie, after all transactions are executed and finalisations applied; formally $H_{\mathrm{r}}$.
\item[transactionsRoot] The Keccak 256-bit hash of the root node of the trie structure populated with each transaction in the transactions list portion of the block; formally $H_{\mathrm{t}}$.
\item[receiptsRoot]\linkdest{receipts_Root_def_words}{} The Keccak 256-bit hash of the root node of the trie structure populated with the receipts of each transaction in the transactions list portion of the block; formally $H_{\mathrm{e}}$.
\item[logsBloom]\linkdest{logs_Bloom_def_words}{} The Bloom filter composed from indexable information (logger address and log topics) contained in each log entry from the receipt of each transaction in the transactions list; formally $H_{\mathrm{b}}$.
\item[difficulty] A scalar value corresponding to the difficulty level of this block. This can be calculated from the previous block's difficulty level and the timestamp; formally $H_{\mathrm{d}}$.
\item[number]\linkdest{block_number_word_def_H_i}{} A scalar value equal to the number of ancestor blocks. The genesis block has a number of zero; formally \hyperlink{block_number_H__i}{$H_{\mathrm{i}}$}.
\item[gasLimit] A scalar value equal to the current limit of gas expenditure per block; formally $H_{\mathrm{l}}$.
\item[gasUsed]\linkdest{block_gas_used_H__g}{}\linkdest{H__g} A scalar value equal to the total gas used in transactions in this block; formally $H_{\mathrm{g}}$.
\item[timestamp]\linkdest{block_timestamp_word_def_H__s}{} A scalar value equal to the reasonable output of Unix's time() at this block's inception; formally \hyperlink{block_timestamp_H__s}{$H_{\mathrm{s}}$}.
\item[extraData]\linkdest{block_extraData_H__x}{} An arbitrary byte array containing data relevant to this block. This must be 32 bytes or fewer; formally $H_{\mathrm{x}}$.
\item[mixHash]\linkdest{mixHash_H__m}{}\linkdest{H__m} A 256-bit hash which, combined with the nonce, proves that a sufficient amount of computation has been carried out on this block; formally $H_{\mathrm{m}}$.
\item[nonce]\linkdest{block_nonce_H__n}{}\linkdest{block_nonce} A 64-bit value which, combined with the mix-hash, proves that a sufficient amount of computation has been carried out on this block; formally \hyperlink{H__n}{$H_{\mathrm{n}}$}.
\end{description}
\linkdest{ommer_block_headers_B__U}{}\linkdest{block_B}{}The other two components in the block are simply a list of ommer block headers (of the same format as above), $B_{\mathbf{U}}$ and a series of the transactions, $B_{\mathbf{T}}$. Formally, we can refer to a block $B$:
\begin{equation}
B \equiv (B_{H}, B_{\mathbf{T}}, B_{\mathbf{U}})
\end{equation}

@@

%Notably $B_{\mathbf{T}}$ does not get serialised into the block by the block preparation function $L_{B}$; it is merely a convenience equivalence.

@@

\linkdest{block_validity}{}We can assert a block's validity if and only if it satisfies several conditions: it must be internally consistent with the ommer and transaction block hashes and the given transactions $B_{\mathbf{T}}$ (as specified in sec \ref{ch:finalisation}), when executed in order on the base state $\boldsymbol{\sigma}$ (derived from the final state of the parent block), result in a new state of the identity $H_{\mathrm{r}}$:
\begin{equation}
\begin{array}[t]{lclc}
\linkdest{new_state_H__r}{}H_{\mathrm{r}} &\equiv& \mathtt{TRIE}(L_S(\Pi(\boldsymbol{\sigma}, B))) & \wedge \\
\linkdest{Ommer_block_hash_H__o}{}H_{\mathrm{o}} &\equiv& \mathtt{KEC}(\mathtt{RLP}(L_H^*(B_{\mathbf{U}}))) & \wedge \\
\linkdest{tx_block_hash_H__t}{}H_{\mathrm{t}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{T}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p (i, L_{T}(B_{\mathbf{T}}[i]))\}) & \wedge \\
\linkdest{Receipts_Root_H__e}{}H_{\mathrm{e}} &\equiv& \mathtt{TRIE}(\{\forall i < \lVert B_{\mathbf{R}} \rVert, i \in \mathbb{N}: &\\&& \quad\quad p(i, B_{\mathbf{R}}[i])\}) & \wedge \\
\linkdest{logs_Bloom_filter_H__b}{}H_{\mathrm{b}} &\equiv& \bigvee_{\mathbf{r} \in B_{\mathbf{R}}} \big( \mathbf{r}_{\mathrm{b}} \big)
\end{array}
\end{equation}
where $p(k, v)$ is simply the pairwise RLP transformation, in this case, the first being the index of the transaction in the block and the second being the transaction receipt:
\begin{equation}
p(k, v) \equiv \big( \mathtt{RLP}(k), \mathtt{RLP}(v) \big)
\end{equation}

Furthermore:
\begin{equation}
\mathtt{TRIE}(L_{S}(\boldsymbol{\sigma})) = {P(B_H)_H}_{\mathrm{r}}
\end{equation}

Thus $\texttt{TRIE}(L_{S}(\boldsymbol{\sigma}))$ is the root node hash of the Merkle Patricia tree structure containing the key-value pairs of the state $\boldsymbol{\sigma}$ with values encoded using RLP, and $P(B_{H})$ is the parent block of $B$, defined directly.

@@

\hypertarget{block_preparation_function_for_RLP_serialization_L__B}{}\linkdest{L__B}\hypertarget{block_preparation_function_for_RLP_serialization_L__H}{}\linkdest{L__B}The function $L_{B}$ and $L_{H}$ are the preparation functions for a block and block header respectively.
We assert the types and order of the structure for when the RLP transformation is required:
\begin{eqnarray}
\quad L_{H}(H) & \equiv & (\begin{array}[t]{l}H_{\mathrm{p}}, H_{\mathrm{o}}, H_{\mathrm{c}}, H_{\mathrm{r}}, H_{\mathrm{t}}, H_{\mathrm{e}}, H_{\mathrm{b}}, H_{\mathrm{d}},\\ H_{\mathrm{i}}, H_{\mathrm{l}}, H_{\mathrm{g}}, H_{\mathrm{s}}, H_{\mathrm{x}}, H_{\mathrm{m}}, H_{\mathrm{n}} \; )\end{array} \\
\quad L_{B}(B) & \equiv & \big( L_{H}(B_{H}), L_{T}^*(B_{\mathbf{T}}), L_{H}^*(\hyperlink{ommer_block_headers_B__U}{B_{\mathbf{U}}}) \big)
\end{eqnarray}

@@

We define $P(B_{H})$ to be the parent block of $B$, formally:
\begin{equation}
P(H) \equiv B': \mathtt{KEC}(\mathtt{RLP}(B'_{H})) = \hyperlink{parent_Hash_H__p_def_words}{H_{\mathrm{p}}}
\end{equation}

\hypertarget{block_number_H__i}{}The block number is the parent's block number incremented by one:
\begin{equation}
H_{\mathrm{i}} \equiv {{P(H)_{H}}_{\mathrm{i}}} + 1
\end{equation}

@@

\hypertarget{block_difficulty_H__d}{}\linkdest{H__d}The canonical difficulty of a block of header $H$ is defined as $D(H)$:
\begin{equation}
D(H) \equiv \begin{dcases}
\mindifficulty & \text{if} \quad H_{\mathrm{i}} = 0\\
\text{max}\!\left(\mindifficulty, {P(H)_{H}}_{\mathrm{d}} + \diffadjustment\times\homesteadmod + \expdiffsymb \right) & \text{otherwise}\\
\end{dcases}
\end{equation}
where:
\begin{equation}
\mindifficulty \equiv 131072
\end{equation}
\begin{equation}
\diffadjustment \equiv \left\lfloor\frac{{P(H)_{H}}_{\mathrm{d}}}{2048}\right\rfloor
\end{equation}
\begin{equation}
\homesteadmod \equiv \text{max}\left( y - \left\lfloor\frac{H_{\mathrm{s}} - {P(H)_{H}}_{\mathrm{s}}}{9}\right\rfloor, -99 \right)
\end{equation}
\begin{equation*}
y \equiv \begin{cases}
1 & \text{if} \, \lVert P(H)_{\mathbf{U}}\rVert = 0 \\
2 & \text{otherwise}
\end{cases}
\end{equation*}
\begin{align}
\expdiffsymb &\equiv \left\lfloor 2^{ \left\lfloor H'_{\mathrm{i}} \div 100000 \right\rfloor - 2 } \right\rfloor \\
H'_{\mathrm{i}} &\equiv \max(H_{\mathrm{i}} - \kappa, 0) \\
\kappa &\equiv \begin{cases} 
  3000000  & \text{if} \quad 4370000 \leqslant H_{\mathrm{i}} < 7280000 \\
  5000000  & \text{if} \quad 7280000 \leqslant H_{\mathrm{i}} < 9200000 \\
  9000000  & \text{if} \quad H_{\mathrm{i}} \geqslant 9200000 \\
\end{cases}
\end{align}

@@

\hypertarget{block_gas_limit_H__l}{}The canonical gas limit $H_{\mathrm{l}}$ of a block of header $H$ must fulfil the relation:
\begin{eqnarray}
& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor \quad \wedge \\
\nonumber& & H_{\mathrm{l}} \geqslant 5000
\end{eqnarray}

\hypertarget{block_timestamp_H__s}{}$H_{\mathrm{s}}$ is the timestamp (in Unix's time()) of block $H$ and must fulfil the relation:
\begin{equation}
H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}}
\end{equation}

@@

\hypertarget{block_header_validity_function}{}Thus we are able to define the block header validity function $V(H)$:
\begin{eqnarray}
V(H) & \equiv &  n \leqslant \frac{2^{256}}{H_{\mathrm{d}}} \wedge m = H_{\mathrm{m}} \quad \wedge \\
\nonumber & & H_{\mathrm{d}} = D(H) \quad \wedge \\
\nonumber& & H_{\mathrm{g}} \le H_{\mathrm{l}}  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} < {P(H)_{H}}_{\mathrm{l}} + \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} > {P(H)_{H}}_{\mathrm{l}} - \left\lfloor\frac{{P(H)_{H}}_{\mathrm{l}}}{1024}\right\rfloor  \quad \wedge \\
\nonumber& & H_{\mathrm{l}} \geqslant 5000  \quad \wedge \\
\nonumber& & H_{\mathrm{s}} > {P(H)_{H}}_{\mathrm{s}} \quad \wedge \\
\nonumber& & H_{\mathrm{i}} = {P(H)_{H}}_{\mathrm{i}} +1 \quad \wedge \\
\nonumber& & \lVert H_{\mathrm{x}} \rVert \le 32
\end{eqnarray}
where $(n, m) = \mathtt{PoW}(H_{\hcancel{n}}, H_{\mathrm{n}}, \mathbf{d})$

@@

\begin{enumerate}
\item The transaction is well-formed RLP, with no additional trailing bytes;
\item the transaction signature is valid;
\item the \hyperlink{transaction_nonce}{transaction nonce} is valid (equivalent to the \hyperlink{account_nonce}{sender account's current nonce});
\item the sender account has no contract code deployed ($\boldsymbol{\sigma}[S(T)]_{\mathrm{c}} = \texttt{KEC}\big( () \big)$).
\item the gas limit is no smaller than the intrinsic gas, $g_0$, used by the transaction; and
\item the sender account balance contains at least the cost, $v_0$, required in up-front payment.
\end{enumerate}

@@

After the message call or contract creation is processed, the refund counter has to be incremented for the accounts that were self-destructed throughout its invocation.
\begin{equation}
	A'_{r} \equiv A_{r} + \sum_{i \in A_{s}} R_{selfdestruct}
\end{equation}

@@

These are used to help define the \hyperlink{Transaction_Receipt}{transaction receipt} and are also used \hyperlink{Upsilon_pow_g2}{later} for state and nonce validation.

@@

$I_{\mathbf{d}}$ evaluates to the empty tuple as there is no input data to this call. $I_{H}$ has no special treatment and is determined from the blockchain.
